An efficient composition of bidirectional programs by tupling and lazy updates

### 2019-10-03 paper/1.introduction.tex
add intro/related/conclusion

In software, there are strong demands for synchronizing the data. In database community it is known as ``view update problem'' and researched for a long time \cite{}.
% A Survey to View Update Problem 
As a solution for this problem, bidirectional transformation is introduced \cite{}. It enables the evaluation of source updating: ``put'' produces the updated source from the original source and the updated view.
% The first paper of bidirectional transformation

There are two types of bidirectional transformation. One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$. A problem of this approach is that it is possible to derive many correct $put$s. To choose one put, hulistics and programmer's intentions are needed. Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$. From a $put$ we can derive an unique $get$.

BiGUL \cite{} is one of put-based bidirectional transformation language. It is fastest bidirectional language in the world now. However, it contains a problem when the BiGUL programs including many compositions. BiGUL does not store the intermediate states when $get$s are re-evaluated so many times. In this case the number of get is quadratic.

\vspace{5mm}

Figure that shows many gets re-evaluation

\vspace{5mm}

@@

To solve this problem and make BiGUL evaluation efficient, we use two techniques. The first technique is tupling, well-known technique for efficiency \cite{}. Its basic idea is to reduce the common computations by combining two evaluations. In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$. Thanks to tupling, we do not lose information like intermediate states.
The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.

The rest of this paper is the following. We introduce a subset of BiGUL, miniBiGUL, and show its syntax, semantics, and properties in Section 2. Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
We apply three optimizations to $pg$ in sequence. The first one is keeping update of BX lazily, and we obtain $cpg$ in Section 4. The second one is lazy evaluation to avoid redundant computations, and we obtain $kpg$ in Section 5. The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs. In Section 7, we show experiemntal results of $pg$, $cpg$, $kpg$, $xpg$ and miniBiGUL.
In Section 8 we discuss about related work, and conclude in Section 9.

@@

\subsection{Contribution}

\begin{itemize}
\item Improvement of evaluation efficiency
  \begin{itemize}
  \item Quadratic to exponential? 
  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
  \end{itemize}
\item Get and put more tight

\item (Tupling introduces some constructions)
  \begin{itemize}
  \item A pair of put and get is self-inverse (involution).
  \end{itemize}
\item careful lazyness is needed in our implementation for efficiency
%  lazy update \& lazy computation are also good optimization for BX.
%\item Reveal a part of relationship between BX and Reversible.
%  \begin{itemize}
%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
%  \item 
%  \item If we can obtain good parameter (dummy) for a BiGUL program,
%    put(s, dummy) can be a complement function.
  \end{itemize}

>>>>

In software, there are strong demands for synchronizing data. In database community this is known as ``view update problem'' and researched for a long time \cite{viewupdate}.
% A Survey to View Update Problem 
As a solution for this problem, bidirectional transformation (BX) is introduced \cite{}. As an example, let us consider a small BX program of $pHead$\footnote{The actual program is shown in the next section.}. BX provides two functions: $get$ and $put$. The following is one example of $pHead$'s evaluation.

\begin{center}
  \includegraphics[width=5cm]{./temp_fig/bxexample1.eps}
\end{center}
%\end{picture}

Here, the original source \texttt{s}, \texttt{[1,2]}, is given by a user. $get$ is a projection: $get$ of $pHead$ picks the first element of the given original source as a view.
%\texttt{get pHead [1,2] = 1}.
After the view \texttt{v} is obtained, the user can modify the view.
In this case, he or she modified the view to \texttt{8} from \texttt{1}.
From this updated view, how can we obtain new updated source? This is ``view update problem.'' For this, BX provides $put$, update on the original source.
In this example, $put$ of $pHead$ constructs a new list \texttt{[8,2]} from the updated view ($v'$) and the original source ($s$).
%\texttt{put pHead [1,2] 8 = [8,2]}

For more various evaluations, we can use composition of programs.
Let us consider another BX example program, three composition of $pHead$: $pHead \circ pHead \circ pHead$.
The following is one example of its evaluation.

\begin{center}
  \includegraphics[width=10cm]{./temp_fig/bxexample2.eps}
\end{center}

In this example, the original source \texttt{s} is \texttt{[[[1,2],[3]],[4]]}.
By repetitious evaluation of $get$s of $pHead$, we can obtain the view $v$, \texttt{1}. To obtain the final updated source \texttt{s'}, we need to evaluate $put$ three times. The first put evaluation is from \texttt{i2} and \texttt{v'} and we can obtain \texttt{i2'}.

For evaluating put-direction, there are two strategies. One is ``not keeping any intermediate states and obtaining them by evaluation when they are needed.'' In this example, ``intermediate states'' are \texttt{i1} and \texttt{i2}.
If we choose this strategy, the number of $get$s will be quadratic. This is a problem, because the evaluation will be slow if the BX programs include many compositions. In this example, two $get$s are required for obtaining \texttt{i2} and one $get$ is required for obtaining \texttt{i1}. In total, three $get$s are required.

The other strategy is ``keeping all intermediate states and using them when they are needed.'' This strategy causes a storage problem. If views includes most information, for evaluating $put$ the needed parts of the original source will be small. It is redundant to keep all intermediate states.

In this work, we choose another strategy ``keeping complements and using them when they are needed.'' Complements are smaller program fragments than the original intermediate states. Therefore this strategy solves the previous two problems.
Readers might already know that it is a hard problem to obtain complements \cite{}. In our work, thanks to the following finding it is not hard:

\vspace{2mm}
In very-well behaved BX programs, 
$put$ can be a complement function for $get$, $get$ can be a complement function for $put$.
\vspace{2mm}

For using complement, we combine $put$ and $get$, and produce new function $pg$ in Section 3.
This is a kind of tupling. Because parts of $put$ and $get$ are doing the same computation, we can make efficient by tupling them.

As an example, let us consider the previous example again: $pHead \circ pHead \circ pHead$.

\begin{center}
  \includegraphics[width=9cm]{./temp_fig/bxexample3.eps}
\end{center}

In the figure, \texttt{c1} and \texttt{c2} are complements. The points in this figure are two. After evaluation of the first $pg$, we do not need to keep the original source \texttt{s}, because all its information is in \texttt{c1} and \texttt{i1}. 
This evaluation looks better than previous two strategies: this does not require repeated evaluation and require smaller storage than the original sources. 
Actually, the simple combined $pg$ is not effective for left associative compositions. To achieve efficient evaluation, we introduce $cpg$, lazy update version of $pg$, in Section 4 and $kpg$, lazy evaluation version of $cpg$, in Section 5. In Section 6 we combine $pg$ and $kpg$ to achieve efficiency for both associative compositions. In Section 7 we compare all approaches including the original BX approach by the evaluation time and memory allocation.

%There are two types of bidirectional transformation. One is get-based bidirectional transformation \cite{} that receives the definition of $get$ and provides $put$. A problem of this approach is that it is possible to derive many correct $put$s. To choose one put, hulistics and programmer's intentions are needed. Another approach is put-based bidirectional transformation \cite{} that received the definition of $put$ and provide $get$. From a $put$ we can derive an unique $get$.

%BiGUL \cite{} is one of put-based bidirectional transformation language. It is fastest bidirectional language in the world now. However, it contains a problem when the BiGUL programs including many compositions. BiGUL does not store the intermediate states when $get$s are re-evaluated so many times. In this case the number of get is quadratic.

%\vspace{5mm}

%Figure that shows many gets re-evaluation

%\vspace{5mm}

@@

%To solve this problem and make BiGUL evaluation efficient, we use two techniques. The first technique is tupling, well-known technique for efficiency \cite{}. Its basic idea is to reduce the common computations by combining two evaluations. In our work, we combines the evaluation of $put$ and $get$, and introduce $pg$. Thanks to tupling, we do not lose information like intermediate states.
%The second technique is lazy update and lazy computation. $pg$ itself includes some inefficiency. For removing the inefficiency we do the update and computation lazily.

%The rest of this paper is the following. We introduce a subset of BiGUL, miniBiGUL, and show its syntax, semantics, and properties in Section 2. Combining the semantics of put and get in Section 2, we construct $pg$ and show that it is self-inverse in Section 3.
%We apply three optimizations to $pg$ in sequence. The first one is keeping update of BX lazily, and we obtain $cpg$ in Section 4. The second one is lazy evaluation to avoid redundant computations, and we obtain $kpg$ in Section 5. The third one is combination of $pg$ and $kpg$ in Section 6, because they are strong for different type of programs. In Section 7, we show experiemntal results of $pg$, $cpg$, $kpg$, $xpg$ and miniBiGUL.
%In Section 8 we discuss about related work, and conclude in Section 9.

@@

\subsubsection*{Contribution of this paper}

We can summarize contribution of this paper as follows.

\begin{itemize}
\item Improvement of evaluation efficiency
  \begin{itemize}
  \item This is both in evaluation time and memory allocation
  \end{itemize}
\item Optimization by tupling
  \begin{itemize}
  \item In BX, as far as authors know optimization by $get$ and $put$ more tight is the first attempt
  \end{itemize}
\end{itemize}

%\begin{itemize}
%\item Improvement of evaluation efficiency
%  \begin{itemize}
%  \item Quadratic to exponential? 
%  \item The introduced extentions (maybe) use more memory than the original approach, but they are faster than the original.
%  \end{itemize}
%\item Get and put more tight

%\item (Tupling introduces some constructions)
%  \begin{itemize}
%  \item A pair of put and get is self-inverse (involution).
%  \end{itemize}
%\item careful lazyness is needed in our implementation for efficiency
%  lazy update \& lazy computation are also good optimization for BX.
%\item Reveal a part of relationship between BX and Reversible.
%  \begin{itemize}
%  \item Very well-behaved (no adaptive-case) \& The same evaluation path $\to$ No garbage Reversible Computation.
%  \item 
%  \item If we can obtain good parameter (dummy) for a BiGUL program,
%    put(s, dummy) can be a complement function.
%  \end{itemize}

### 2019-10-03 paper/8.related-work.tex
add intro/related/conclusion

\begin{itemize}
    %\item BiGUL papers (PEPM, POPL, tutorial?)
    %\item the first BX paper
    %\item get-based BX
    %\item put-based BX?
    \item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
    \item optimization techniques : tupling, lazy computation, fusion transformation?
    \item (view update problem paper)
\end{itemize}

>>>>

\subsubsection*{BX langauges and implementations}

There are many BX languages \cite{josh_pepm} \cite{BXtend} \cite{eMoflon} \cite{EVL+Strace} \cite{JTL} \cite{NMF} (memo: BXtend, eMoflon, EVL+Strace, JTL, NMF. SDMLib should be added?) based on several application scenarios and BX researches \cite{} (add papers for lenses). They mostly focus on the theory side, such as semantics and correctness, and do not focus on efficiency so much. For practical implementation of BX languages Anjorin et. al. introduces the first benchmark for BX \cite{BXcomp}. We can say this paper will be first attempt to improve efficiency of BX composition evaluation. In this paper we focus on BiGUL \cite{josh_pepm} \cite{josh_popl}, its implementation uses ``not keeping any intermediate states and obtaining them by evaluation when they are needed'' strategy. Other BX languages might have the same problem, and our approach can be applicable.

\subsubsection*{Optimization techniques}

In this paper, we use several optimization techniques: tupling, lazy update, and lazy computation (\textcolor{red}{also fusion?}). Tupling \cite{tupling} is an optimization technique by combining several computations.
Lazy update .. delta lenses?

Lazy computation \cite{} is also an old but effective widly-used technique.

\begin{itemize}
\item This part looks just short introduction of  techniques .. 
\end{itemize}

%\begin{itemize}
    %\item BiGUL papers (PEPM, POPL, tutorial?)
    %\item the first BX paper
    %\item get-based BX
    %\item put-based BX?
%    \item implementation of BX, they might have the same problem, this paper will be first attempt, semantics and correctness is considered more, but efficiency is  lacked.
%    \item optimization techniques : tupling, lazy computation, fusion transformation?
%    \item (view update problem paper)
%\end{itemize}

### 2019-10-03 paper/main.tex
add intro/related/conclusion

\begin{document}
%
\title{An efficient composition of bidirectional programs
  by tupling and lazy updates 
%  tupling
%  accumulating updates = lazy updates
%  based on the idea of reversible computing  
  % \thanks{Supported by organization x.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The abstract should briefly summarize the contents of the paper in
15--250 words.

\keywords{First keyword  \and Second keyword \and Another keyword.}
\end{abstract}
%
%
%

@@

Schedule
\begin{itemize}

\item Introduction (story), Contribution, positions, examples, beginning of September?
\item zero draft : mid of September

\item deadline of PEPM2020 (October 18th)
\item ESOP (October 24th)
\item FLOPS (the middle of November??)
\item Need to change the style file of latex..
\item []
\end{itemize}

@@

% From sample code
% 
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{ref_article1}
Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

\bibitem{ref_lncs_1}
Author, F., Author, S.: Title of a proceedings paper. In: Editor,
F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
Springer, Heidelberg (2016). \doi{10.10007/1234567890}

\bibitem{ref_book1}
Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
Location (1999)

\bibitem{ref_proc1}
Author, A.-B.: Contribution title. In: 9th International Proceedings
on Proceedings, pp. 1--2. Publisher, Location (2010)

\bibitem{ref_url1}
LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
Oct 2017
\end{thebibliography}

>>>>

\begin{document}
%
\title{An efficient composition of bidirectional programs
  by tupling and lazy updates 
%  tupling
%  accumulating updates = lazy updates
%  based on the idea of reversible computing  
  % \thanks{Supported by organization x.}
}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{First Author\inst{1}\orcidID{0000-1111-2222-3333} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
%The abstract should briefly summarize the contents of the paper in
%15--250 words.
  Bidirectional transformation (BX) is a solution of view update problems and widely used for synchronizing data. Its semantics and correctness are researched well, but the efficiency and optimization are not considered well. In this work we focus on the evaluation of composition, because it includes some problems of evaluation time efficiency and memory allocation. To solve these problems we make $put$ and $get$ more tight, based on the idea of tupling. %42
  Because simple tupled result includes some redundancies for left associative compositions, we apply two optimization techniques: lazy update and lazy evaluation. We show some experimental results that our optimized approach is faster than the original approach that is used in an actual BX language.

\keywords{Bidirectional transformation  \and Tupling \and Optimization \and ??.}
\end{abstract}
%
%
%

@@

%Schedule
\begin{itemize}
%\item Introduction (story), Contribution, positions, examples, beginning of September?
%\item zero draft : mid of September  
%\item deadline of PEPM2020 (October 18th)
%\item ESOP (October 24th)
\item FLOPS 2020, 15 November 2019 (AoE): Abstract submission
\item 22 November 2019 (AoE): Submission deadline
\item keywords, ORCID id, introduction, contribution, future work, related work 
\end{itemize}

@@

\textcolor{red}{TODO: FIX reference, order, contents ..}

% From sample code
% 
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
% \bibliographystyle{splncs04}
% \bibliography{mybibliography}
%
\begin{thebibliography}{8}
\bibitem{viewupdate}
F. Bancilhon and N. Spyratos. ``Update semantics of relational views''. ACM Transactions on Database Systems, 6(4):557--575, 1981.

\bibitem{BXcomp}
Anthony Anjorin, Thomas Buchmann, Bernhard Westfechtel, Zinovy Diskin, Hsiang-Shang Ko, Romina Eramo, Georg Hinkel, Leila Samimi-Dehkordi, and Albert Zündorf
``Benchmarking bidirectional transformations: Theory, implementation, application, and assessment'', Software and Systems Modeling, ??, 2019

\bibitem{tupling}
  M. Fokkinga. ``Tupling and mutumorphisms''. Squiggolist, 1(4), 1989.

\bibitem{josh_pepm}
  Hsiang-Shang Ko, Tao Zan, Zhenjiang Hu, BiGUL: A Formally Verified Core Language for Putback-Based Bidirectional Programming, ACM SIGPLAN 2016 Workshop on Partial Evaluation and Program Manipulation (PEPM 2016), pp.61--72, 2016.

\bibitem{josh_popl}
  Hsiang-Shang Ko, Zhenjiang Hu, An Axiomatic Basis for Bidirectional Programming , 45th ACM SIGPLAN Symposium on Principles of Programming Languages (POPL 2018), 2018.

\bibitem{BXtend}
  Thomas Buchmann. BXtend -- a framework for (bidirectional) model transformations. In Slimane Hamoudi,Luis Ferreira Pires, and Bran Selic, editors, Proceedings
of the 6th International Conference on Model-Driven Engineering and Software Development - Volume 1: MODELSWARD (MODELSWARD 2018), pages 336--345, 2018.

\bibitem{eMoflon}
  Erhan Leblebici, Anthony Anjorin, and Andy Sch"{u}rr. Developing eMoflon with eMoflon. In Davide Di Ruscio and Daniel Varr´o, editors, Theory and Practice of Model Transformations - 7th International Conference, ICMT 2014, Held as Part of STAF 2014, York, UK, 2014. Proceedings, volume 8568 of Lecture Notes in
Computer Science, pages 138–145. Springer, 2014.

\bibitem{EVL+Strace}
Leila Samimi-Dehkordi, Bahman Zamani, and Shekoufeh Kolahdouz-Rahimi. EVL+Strace: A novel bidirectional transformation approach. Information and Software
Technology, 100:47--72, 2018.

\bibitem{JTL}
Antonio Cicchetti, Davide Di Ruscio, Romina Eramo, and Alfonso Pierantonio. JTL: A bidirectional and change propagating transformation language. In Brian Malloy, Steffen Staab, and Mark van den Brand, editors, Proceedings of the Third International Conference on Software Language Engineering (SLE 2010), volume 6563 of Lecture Notes of Computer Science, pages
183--202, Springer-Verlag.

\bibitem{NMF}
Georg Hinkel and Erik Burger. Change propagation and bidirectionality in internal transformation DSLs. Soft-ware and Systems Modeling, 18(1):249--278, 2019.

\bibitem{SDMLib}
  SDMLib paper?

\bibitem{lenses}
  Lense papers?

\bibitem{tutorial}
  Tutorial paper of BiGUL

%\bibitem{ref_article1}
%Author, F.: Article title. Journal \textbf{2}(5), 99--110 (2016)

%\bibitem{ref_lncs_1}
%Author, F., Author, S.: Title of a proceedings paper. In: Editor,
%F., Editor, S. (eds.) CONFERENCE 2016, LNCS, vol. 9999, pp. 1--13.
%Springer, Heidelberg (2016). \doi{10.10007/1234567890}

%\bibitem{ref_book1}
%Author, F., Author, S., Author, T.: Book title. 2nd edn. Publisher,
%Location (1999)

%\bibitem{ref_proc1}
%Author, A.-B.: Contribution title. In: 9th International Proceedings
%on Proceedings, pp. 1--2. Publisher, Location (2010)

%\bibitem{ref_url1}
%LNCS Homepage, \url{http://www.springer.com/lncs}. Last accessed 4
%Oct 2017
\end{thebibliography}
